Lesson 3
========================================================

***

### What to Do First?
Notes:

***

### Pseudo-Facebook User Data
Notes:

```{r Pseudo-Facebook User Data}
setwd('C:\\Users\\ransf\\OneDrive\\Documents\\GitHub\\Data-Analysis-With-R\\Lesson3')
pf <- read.csv("pseudo_facebook.tsv",sep = "\t")
names(pf)
```

***

### Histogram of Users' Birthdays
Notes:

```{r Histogram of Users\' Birthdays}
install.packages("ggplot2")
library(ggplot2)
qplot(x=dob_day, data=pf) +
  scale_x_discrete(breaks=1:31)
```

***

#### What are some things that you notice about this histogram?
Response:

***

### Moira's Investigation
Notes:Moira investigation was about taking a set of Facebook users and performing a survey on the perceived audience size of their posts.  Participants were asked what percentage of their facebook users actually saw the post.  Link to the paper is [here](http://hci.stanford.edu/publications/2013/invisibleaudience/invisibleaudience.pdf)
***

### Estimating Your Audience Size
Notes:

***

#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response: When I got engaged.

#### How many of your friends do you think saw that post?
Response:400

#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is?
Response:40%

***

### Perceived Audience Size
Notes: Based on Moira's Investigation, people underestimate their audience by about 3/4th.  This means that their guessed audience size is only about a fourth of the actual audience size.

***
### Faceting
Notes: Faceting is the idea of splitting up graphs into multiple graphs based on a particular variable. **facet_wrap** is best used when generating plots for one variable. **facet_grid** is best used when you have two or more variables.

```{r Faceting}
qplot(x=dob_day,data = pf) +
  scale_x_discrete(breaks= 1:31) +
  facet_wrap(~dob_month,ncol=3)
```

#### Let’s take another look at our plot. What stands out to you here?
Response:

***

### Be Skeptical - Outliers and Anomalies
Notes:

***

### Moira's Outlier
Notes:
#### Which case do you think applies to Moira’s outlier?
Response:

***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}
qplot(x=friend_count,data=pf)
```

#### How is this plot similar to Moira's first plot?
Response:
The data
***

### Limiting the Axes
Notes:

```{r Limiting the Axes}
qplot(x=friend_count,data=pf) +
  scale_x_continuous(limits=c(0,1000))
```

### Exploring with Bin Width
Notes:

***

### Adjusting the Bin Width
Notes:Here we adjust the binwidth to get rid of the warning message in the console.  This is done by setting the *binwidth* parameter in the **qplot** function.  You can also adjust the breakpoints in the graph by using the *breaks* parameter in the **scale_x_continuous** function.  More information about scales function can be found [here](http://docs.ggplot2.org/current/scale_continuous.html)
```{r Adjusting the Bin Width}
qplot(x=friend_count,data = pf,binwidth = 25) + scale_x_continuous(limits = c(0,1000), breaks = seq(0,1000,50))
```

### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count, data = pf, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```

***

### Omitting NA Values
Notes:
To omit the NA values in our gender column, we modify the data to take in the subset of data that does not have a NA value for gender. You can also perform this using the **na.omit()** function, but this may be too aggressive (It may omit genuine NA values).
```{r Omitting NA Values}
qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),breaks = seq(0, 1000, 50)) + 
  facet_wrap(~gender)
```

***

### Statistics 'by' Gender
Notes: To get the number of females and males can given using the **table** function.  To get the average number of friends by gender, you can use the **by** function. The **by** functions takes a parameter, how you want to split, and a function to run on the first two parameters.  We will use the **summary** function in this example

```{r Statistics \'by\' Gender}
table(pf$gender)
by(pf$friend_count,pf$gender,summary)

```

#### Who on average has more friends: men or women?
Response: Women

#### What's the difference between the median friend count for women and men?
Response: 22

#### Why would the median be a better measure than the mean?
Response: Because the median is more robust and is not affected by outliers

***

### Tenure
Notes: The *I()* function stands for 'as is' which tells qplot to use them as colors.

```{r Tenure}
#Plot split based on days
qplot(x=tenure,data=pf, binwidth = 30,
      color = I('black'), fill = I('#099DD9'))

```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}
#Plot split based on years
qplot(x=tenure/365,data=pf, binwidth = 1,
      color = I('black'), fill = I('#099DD9'))
```

***

### Labeling Plots
Notes:

```{r Labeling Plots}
qplot(x=tenure/365,data=pf, binwidth = 0.25,
      xlab = 'Number of years using Facebook',
      ylab = 'Numbers of users in sample',
      color = I('black'), fill = I('#099DD9')) +
  scale_x_continuous(limits = c(0,7), breaks = seq(1,7,1 ))
```

***

### User Ages
Notes:

```{r User Ages}
qplot(x = age, data = pf, binwidth = 5,
      xlab = "Age of users within a specific age range",
      ylab = "Number of users within each age range",
      color = I('black'), fill = I('#099DD9')) + 
  scale_x_continuous(limits = c(10,100), breaks = seq(0,100,10))
```

#### What do you notice?
Response:

***

### The Spread of Memes
Notes:

***

### Lada's Money Bag Meme
Notes:

***

### Transforming Data
Notes:
```{r Transforming Data}
qplot(x = friend_count, data = pf)

summary(pf$friend_count)

summary(log10(pf$friend_count + 1))

summary(sqrt(pf$friend_count))

```
***

### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}

```

***


### Frequency Polygons

```{r Frequency Polygons}

```

***

### Likes on the Web
Notes:

```{r Likes on the Web}

```


***

### Box Plots
Notes:

```{r Box Plots}

```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}

```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}

```

#### On average, who initiated more friendships in our sample: men or women?
Response:
#### Write about some ways that you can verify your answer.
Response:
```{r Friend Requests by Gender}

```

Response:

***

### Getting Logical
Notes:

```{r Getting Logical}

```

Response:

***

### Analyzing One Variable
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!