Lesson 5
========================================================

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
setwd('C:\\Users\\ransf\\OneDrive\\Documents\\GitHub\\Data-Analysis-With-R\\Lesson3')
pf <- read.csv("pseudo_facebook.tsv",sep = "\t")
library(ggplot2)
ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) + geom_boxplot() +
  stat_summary(fun.y = mean,geom = "point", shape = 4)

ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(gender))) + geom_line(aes(color = gender), stat = 'summary', fun.y = median)

pf.fc_by_age_gender <- subset(pf, !is.na(gender)) %>% group_by(age,gender) %>% summarise(mean_friend_count=mean(friend_count),median_friend_count=median(friend_count), n=n())
  
```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
ggplot(pf.fc_by_age_gender, aes(x=age,y=median_friend_count)) + geom_line(aes(color=gender))
```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:

```{r}
pf.fc_by_age_gender.wide <- subset(pf.fc_by_age_gender, select = c(age,gender,median_friend_count)) %>% spread(gender,median_friend_count) %>% mutate(ratio = female/male)
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
ggplot(pf.fc_by_age_gender.wide, aes(x=age,y=ratio)) + geom_line() +
  geom_hline(linetype = 2, yintercept=1)
```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
pf <- pf %>% mutate(year_joined = as.integer(2014 - (tenure/365)))
table(pf$year_joined)



```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
pf$year_joined.bucket <- cut(pf$year_joined, c(2004,2009,2011,2012,2014))

table(pf$year_joined.bucket)
```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(year_joined.bucket))) + geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median)
```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(year_joined.bucket))) + geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean) +
  geom_line(stat = 'summary', fun.y=mean, linetype =2 )
```

***

### Friending Rate
Notes:

```{r Friending Rate}
pf <- subset(pf,tenure >=1) %>% mutate(friend_rate = friend_count/tenure)
summary(pf$friend_rate)
```

***

### Friendships Initiated
Notes:

What is the median friend rate? 0.2205

What is the maximum friend rate? 417

```{r Friendships Initiated}
ggplot(aes(x = tenure, y = friendships_initiated/tenure),
       data = subset(pf, !is.na(year_joined.bucket) & tenure >=1)) + geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean)
```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
setwd('C:\\Users\\ransf\\OneDrive\\Documents\\GitHub\\Data-Analysis-With-R\\Lesson5')
yo <- read.csv('yogurt.csv')
yo$id <- factor(yo$id)
str(yo)
```

***

### Number of Purchases
Notes:

```{r Number of Purchases}

```

***

### Prices over Time
Notes:

```{r Prices over Time}

```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}

```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

